---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(readr)
library(janitor)
library(stringr)
library(here)
```

#read data
```{r}
raw_data <- read_rds("raw_data/decathlon.rds")
```

# 1.1.1 - Cleaning
```{r}
cleaned_data <- rownames_to_column(raw_data, var = "athlete") 

names(cleaned_data) <- c("athlete", "run_100m", "long_jump", "shot_put", "high_jump", "run_400m", "hurdles_110m", "discus", "pole_vault", "javeline", "run_1500m", "rank", "points", "competition")

cleaned_data
```

```{r}
decathlon_data <- cleaned_data %>%
  mutate(athlete = str_to_lower(athlete)) %>%
  pivot_longer(cols = c(run_100m:run_1500m),
  names_to = "event",
  values_to = "result")
decathlon_data
```

#1 Who had the longest long jump seen in the data?
```{r}
decathlon_data %>%
  filter(event == "long_jump") %>%
  arrange(desc(result))
```


#2 What was the average 100m time in each competition?
```{r}
decathlon_data %>%
  filter(event == "run_100m") %>%
  group_by(competition) %>%
  summarise(mean(result)) 
```




# Who had the highest total points across both competitions?
```{r}
decathlon_data %>%
  group_by(competition) %>%
  summarise(max(points)) 
```



# What was the shot-put scores for the top three competitors in each competition?
```{r}
decathlon_data %>%
  filter(event == "shot_put") %>%
  arrange(desc(result)) %>%
  group_by(competition) %>%
  top_n(3)
  
```


# What was the average points for competitors who ran the 400m in less than 50 seconds vs. those than ran 400m in more than 50 seconds?
```{r}

```









